Disclaimer: This document represents notes taken from Advanced Data Analysis: STAT 527 and is not
work of my own.

# Learning objectives
1. Identify a function or operation and describe it's use.
2. Apply functions and operations to achieve a specific result.
3. Predict answers of calculations written in R.
4. Use R's functions to get help and numerically summarize data.
5. Apply ggplot() to organize and reveal patterns visually.
6. Explain what each plotting option does.

# R building blocks
### Basic arithmetic operations
The following examples demonstrates basic functionality of R as well as various data types
available in R.

```{r}
# Arithmetic
2 * 10
1 + 2

# Order of operations
1 + 5 * 10
(1 + 5) * 10

# Exponents
2^10
9^(1/2)
```

### Vectors
A vector is a set of numbers similar to the columns in a spreadsheet. In R these can be ordered and indexed.
```{r}
# Vector
c(1, 2, 3, 4)
c(1:5, 10)

# Using seq to create a sequence
seq(from=1, to=10, by=2)
seq(1, 10, by=2)
seq(1, 10, length=11)
seq(1, 10)

# Creating sequences using :
1:5
```

### Assignment and variables
Data type (i.e. sequences, integers, strings, data.frames) can all be assigned to variables using the `<-` operator. Variables
created in this way are stored in memory and can be operated on and referenced by calling the variable name.
```{r}
# Assign a vector to variable a
a <- 1:5
a

b <- seq(15, 3, length=5)
b

c <- a * b
```

### Basic functions
The R-base package has many functions available for performing routine takes such as computing the mean of a vector, summing a vector
as well as several other tasks. Help documentation for a particular function can be viewed by using putting a _?_ at the beginning
of the function name, `?sum`, `?data sets`.
```{r}
a
sum(a)
mean(a)
sd(a)
prod(a)
var(a)
min(a)
max(a)
median(a)
range(a)
```

### Extracting subsets
Lists, vectors, data frames and matrices can all be subseted using various techniques.
```{r}
# Create a vector ranging from 0 to 100 by 10
a <- seq(0, 100, by=10)

# Index/subset the first element of the vector
a[1]

# Index the first 3 elements of the vector
a[1:3]

# Index the first and fourth elements
a[c(1,4)]

# Reassign the value of the first element
a[1] <- 7

# Evaluate elements of a vector
a > 50

# Subset vector based if the evaluate to TRUE
a[a > 50]

# Similarly, subset elements based on conditionals
which(a > 50)

# Negate evaluation
!(a > 50)

# Select all elements NOT greater than 50
a[!(a > 50)]
```

### Comparison functions
Comparisons return a boolean type object when evaluated.

```{r}
a

# Extract elements where expression evaluates to TRUE
a[(a == 55)]
a[(a != 55)]
a[(a > 50)]
a[(a < 50)]
a[(a >= 50)]
a[(a <= 50)]

# Set operations
c(10, 14, 40, 60, 99) %in% a
```

### Boolean operators
Compare TRUE/FALSE values and return TRUE/FALSE values.

```{r}
a

# Subset values within a certain range using and
a[(a > 50) & (a <=90)]

# Subset either values using or
a[(a < 50) | (a > 100)]
a[(a < 50) | !(a > 100)]
a[(a >= 50) & !(a <= 90)]
```

### Missing values
The **NA** (not available) means the value is missing. Any calculation involving **NA** will return a **NA** by default.

```{r}
NA + 8
3 * NA
mean(c(1, 2, NA))

# Some functions have the ability to ignore NA
mean(c(NA, 1, 2), na.rm=TRUE)
sum(c(NA, 1, 2))
sum(c(NA, 1, 2), na.rm=TRUE)

# Evaluating NA
a <- c(NA, 1:5, NA)
a
is.na(a)
!is.na(a)

# Subset all values that are not NA
a[!is.na(a)]
```

# Plotting with ggplot2
This section is intended as an introduction to `ggplot2` and some of its capabilities. As a basic introduction, it requires a `data.frame` object as input, and then you define plot layers that stack on top of each other, and each layer has visual/text elements that are mapped to aesthetics (colors, size, opacity). In this way, a simple set of commands can be combined to produce extremely informative displays.

In the example that follows, we consider a data set `mpg` consisting of fuel economy data from 1999 and 2008 for 38 popular models of car.

```{r}
# Install ggplot2 if not already
if(!require(ggplot2)){install.packages("ggplot2")}

# Load the ggplot2 package
library(ggplot2)

# The mpg data set should be already loaded into our R environment
# lets take a look at the top few rows of data using head()
head(mpg)

# Inspect the data type of each column
str(mpg)

# Compute summary statistics on each column/data type
summary(mpg)

# Plot hwy mpg against engine displacement adding titles and axis labels
ggplot(mpg) + geom_point(aes(x=displ, y=hwy)) + labs(title="Highway MPG vs Displacement") +
  xlab("Engine displacement (liters)") + ylab("Highway MPG")
```

The `ggplot()` function sets up the data frame to be used while the `geom_point()` function specifies the type of plot (i.e. geom) to use. The `aes()` function inside `geom_point()` is used to map plot attributes (i.e. $x$ and $y$ values) to columns within the data frame. We can map other values to plot attributes using `aes()` for example, we can map the color of each point to the class of vehicle.

```{r}
ggplot(mpg) + geom_point(aes(x=displ, y=hwy, color=class)) + labs(title="Highway MPG vs Displacement") +
  xlab("Engine displacement (liters)") + ylab("Highway MPG")
```
We can go a step further and map the size of each point to the number of cylinders and the shape of each point to the drive type (i.e. number of gears). In this way, we can encode a large amount of information (dimensions) into each plot (although this is not always desirable).

```{r}
ggplot(mpg) + geom_point(aes(x=displ, y=hwy, color=class, shape=drv, size=cyl)) +
  labs(title="Highway MGP vs Displacement") + xlab("Engine displacement (liters") +
  ylab("Highway MPG")
```

### Faceting
Small multiple plots (i.e. faceting, trellis chart, lattice chart, grid chart, panel chart) is a series or grid of small similar
graphics or charts. These are useful when we want to stratify and compare groups of data to one another. Typically facets are formed on different subsets of data and are really useful for exploring conditional relationships between data.

```{r}
# Start by creating a basic scatterplot
p <- ggplot(mpg) + geom_point(aes(x=displ, y=hwy))

p1 <- p + facet_grid(. ~ cyl)   # Columns are cyl categories
p2 <- p + facet_grid(drv ~ .)   # Rows are drv categories
p3 <- p + facet_grid(drv ~ cyl) # Rows are drv categories, columns are cyl categories
p4 <- p + facet_wrap(~ class)   # Wrap plots by class category

# Plot all plots in one figure
library(gridExtra)
grid.arrange(p1, p2, p3, p4, ncol=2)
```

# Improving plots
Sometimes plots can be enhanced by adding noise. One example is when plotting categorical data or large numbers of data. For example, suppose we are plotting large numbers of categorical data, because categorical values take on finite values, we can expect
much of the data points to overlap and obfuscate one another. One solution to this is to add just enough noise that the data points separate while still maintaining their categorical relationships. 

```{r}
# Obfuscated data
ggplot(mpg) + geom_point(aes(x=cty, y=hwy))
```
In the above plot, some of the data overlap. We can avoid this issue by adding jitter to each point and changing the transparency (alpha) of each point so that hidden data points are revealed.

```{r}
ggplot(mpg) + geom_point(aes(x=cty, y=hwy), position = "jitter", alpha=0.5)
```

We can also improve plots by rearranging the order in which data appear.

```{r}
ggplot(mpg) + geom_point(aes(x=class, y=hwy))
```
Suppose we want to reorder each class ascendingly based on highway MPG

```{r}
ggplot(mpg) + geom_point(aes(x=reorder(class, hwy), y=hwy))
```

Great, we reordered the vehicle class by highway MPG but a lot of the data are still obfuscated. Lets try adding jitter and
changing transparency.

```{r}
ggplot(mpg) + geom_point(aes(x=reorder(class, hwy), y=hwy), position="jitter", alpha=0.5)
```

Suppose were are more interested in comparing the distribution of highway MPG between vehicle classes. Box plots would be more appropriate

```{r}
ggplot(mpg) + geom_boxplot(aes(x=reorder(class, hwy), y=hwy))
```

That's good, but we would like to see how the data points are distributed within the quartiles, lets add a `geom_point()` layer to show this

```{r}
# Specify how much jitter you would like to add
ggplot(mpg) + geom_boxplot(aes(x=reorder(class, hwy), y=hwy)) +
  geom_point(aes(x=reorder(class, hwy), hwy), alpha=0.5, position=position_jitter(width=0.1)) +
  labs(title="Vehicle class vs highway MPG") + xlab("Vehicle class") + ylab("Highway MPG")
```

The the above examples we reordered the vehicle class based on mean highway MPG (mean is used by default), however we can just as easily reorder based on other statistics by specifying what `FUN` to use

```{r}
# Order vehicle class by median MPG
ggplot(mpg) + geom_boxplot(aes(x=reorder(class, hwy, FUN=median), y=hwy), alpha=0.5) +
  geom_point(aes(x=reorder(class, hwy), hwy), alpha=0.5, position=position_jitter(width=0.1)) +
  labs(title="Ordered by median MGP")
```

```{r}
# Order vehicle class by the MPG standard deviation
ggplot(mpg) + geom_boxplot(aes(x=reorder(class, hwy, FUN=sd), y=hwy), alpha=0.5) +
  geom_point(aes(x=reorder(class, hwy), hwy), alpha=0.5, position=position_jitter(width=0.1)) +
  labs(title="Ordered by MPG standard deviation")
```